Age-related changes in plant growth and traits

### 2015-05-20 counts.Rmd
Rename downstream variables

plot(seedset ~ FinalWeight,data=subset(SummaryInd,seedset>0),
          col=col.spp[as.factor(species)],
          pch=16,log="xy",xlab="plant weight",ylab="seed set") 
plot(seedset ~ RA,data=subset(SummaryInd,seedset>0),
          col=col.spp[as.factor(species)],
          pch=16,log="xy",ylab="")
plot.new()

>>>>

plot(seedset ~ total_weight,data=subset(SummaryInd,seedset>0),
          col=col.spp[as.factor(species)],
          pch=16,log="xy",xlab="plant weight",ylab="seed set") 
plot(seedset ~ RA,data=subset(SummaryInd,seedset>0),
          col=col.spp[as.factor(species)],
          pch=16,log="xy",ylab="")
plot.new()

### 2015-05-20 leaflifespan.Rmd
Rename downstream variables

#adding variables to look at change across years, RGR
SummaryInd$change_shoot_diam <- SummaryInd$d_end - SummaryInd$d_start
SummaryInd$change_basal_diam <- SummaryInd$FinalBasalDiam - SummaryInd$StartBasalDiam
SummaryInd$change_shoot_area <- (3.14*((SummaryInd$d_end/2)^2)) - (3.14*((SummaryInd$d_start/2)^2))
SummaryInd$change_basal_area <- (3.14*((SummaryInd$FinalBasalDiam/2)^2)) - (3.14*((SummaryInd$StartBasalDiam/2)^2))
SummaryInd$lvs_end_total <- SummaryInd$lvs_end + SummaryInd$lvs_new
SummaryInd$RGR <- log(SummaryInd$FinalWeight)-log(SummaryInd$FinalWeight-SummaryInd$GrowthInv)
```

@@

#investment summary by species, age
investment_summary <- investment %>%
  group_by(species, age) %>%
  summarise_each(funs(mean, se, length), GrowthInv, FinalWeight, FinalBasalDiamAv, TotalInvestment, RA)

@@

```{r, echo=FALSE,results='hide', message=FALSE}
plot(total_leaf_weight~FinalWeight,SummaryInd, log="xy",col=col.age[as.factor(age)], pch=16, cex=0.8)
legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
mod <-lm(FinalWeight~total_leaf_weight+age+species,SummaryInd)
summary(mod)

plot_LeafWeight <- function(spp) {
  plot(total_leaf_weight~FinalWeight,data=subset(SummaryInd,species==spp), xlab="total weight (mg)", ylab ="leaf weight (mg)", log="xy",col=col.age[as.factor(age)], pch=16,     cex=0.8,main=labels.spp[spp])
  legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
  mod <-lm(FinalWeight~total_leaf_weight+age,data=subset(SummaryInd,species==spp))
  abline(mod)
  summary(mod)
}

@@

#plot looking at leaf area versus plant weight
#USE THIS PLOT IN LAB MEETING - CLEARLY SHOWS DECLINE IN INTERCEPT WITH AGE
plot(leafArea~FinalWeight,SummaryInd, log="xy",col=col.age[as.factor(age)], pch=16, cex=0.8)
legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
mod <-lm(FinalWeight~leafArea+age+species,SummaryInd)
summary(mod)

plot_LeafArea <- function(spp) {
  plot(leafArea~FinalWeight,data=subset(SummaryInd,species==spp), xlab="total weight (mg)", ylab ="leaf weight (mg)", log="xy",col=col.age[as.factor(age)], pch=16,     cex=0.8,main=labels.spp[spp])
  legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
  mod <-lm(FinalWeight~leafArea+age,data=subset(SummaryInd,species==spp))
  abline(mod)
  summary(mod)
}

@@

#plots looking at shoot extension versus plant size
plot(new_length~FinalWeight,SummaryInd, log="xy",col=col.age[as.factor(age)], pch=16, cex=0.8)
legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
mod <-lm(FinalWeight~new_length+age+species,SummaryInd)
summary(mod)

plot_LeaderExtension <- function(spp) {
  plot(new_length~FinalWeight,data=subset(SummaryInd,species==spp), xlab="total weight (mg)", ylab ="leaf weight (mg)", log="xy",col=col.age[as.factor(age)], pch=16,     cex=0.8,main=labels.spp[spp])
  legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
  mod <-lm(FinalWeight~new_length+age,data=subset(SummaryInd,species==spp))
  abline(mod)
  summary(mod)
}

>>>>

#adding variables to look at change across years, RGR
SummaryInd$change_shoot_diam <- SummaryInd$d_end - SummaryInd$d_start
SummaryInd$change_basal_diam <- SummaryInd$growth_stem_diam
SummaryInd$change_shoot_area <- (3.14*((SummaryInd$d_end/2)^2)) - (3.14*((SummaryInd$d_start/2)^2))
SummaryInd$change_basal_area <- SummaryInd$growth_stem_area
SummaryInd$lvs_end_total <- SummaryInd$lvs_end + SummaryInd$lvs_new
SummaryInd$RGR <- log(SummaryInd$total_weight)-log(SummaryInd$total_weight-SummaryInd$GrowthInv)
```

@@

#investment summary by species, age
investment_summary <- investment %>%
  group_by(species, age) %>%
  summarise_each(funs(mean, se, length), GrowthInv, total_weight, dia, TotalInvestment, RA)

@@

```{r, echo=FALSE,results='hide', message=FALSE}
plot(total_leaf_weight~total_weight,SummaryInd, log="xy",col=col.age[as.factor(age)], pch=16, cex=0.8)
legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
mod <-lm(total_weight~total_leaf_weight+age+species,SummaryInd)
summary(mod)

plot_LeafWeight <- function(spp) {
  plot(total_leaf_weight~total_weight,data=subset(SummaryInd,species==spp), xlab="total weight (mg)", ylab ="leaf weight (mg)", log="xy",col=col.age[as.factor(age)], pch=16,     cex=0.8,main=labels.spp[spp])
  legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
  mod <-lm(total_weight~total_leaf_weight+age,data=subset(SummaryInd,species==spp))
  abline(mod)
  summary(mod)
}

@@

#plot looking at leaf area versus plant weight
#USE THIS PLOT IN LAB MEETING - CLEARLY SHOWS DECLINE IN INTERCEPT WITH AGE
plot(leafArea~total_weight,SummaryInd, log="xy",col=col.age[as.factor(age)], pch=16, cex=0.8)
legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
mod <-lm(total_weight~leafArea+age+species,SummaryInd)
summary(mod)

plot_LeafArea <- function(spp) {
  plot(leafArea~total_weight,data=subset(SummaryInd,species==spp), xlab="total weight (mg)", ylab ="leaf weight (mg)", log="xy",col=col.age[as.factor(age)], pch=16,     cex=0.8,main=labels.spp[spp])
  legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
  mod <-lm(total_weight~leafArea+age,data=subset(SummaryInd,species==spp))
  abline(mod)
  summary(mod)
}

@@

#plots looking at shoot extension versus plant size
plot(new_length~total_weight,SummaryInd, log="xy",col=col.age[as.factor(age)], pch=16, cex=0.8)
legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
mod <-lm(total_weight~new_length+age+species,SummaryInd)
summary(mod)

plot_LeaderExtension <- function(spp) {
  plot(new_length~total_weight,data=subset(SummaryInd,species==spp), xlab="total weight (mg)", ylab ="leaf weight (mg)", log="xy",col=col.age[as.factor(age)], pch=16,     cex=0.8,main=labels.spp[spp])
  legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
  mod <-lm(total_weight~new_length+age,data=subset(SummaryInd,species==spp))
  abline(mod)
  summary(mod)
}

### 2015-05-20 results_RAschedules.Rmd
Rename downstream variables

#RV curves
###Classic RV curves - reproductive investment versus weight
- These are mostly reasonably nice - of course need to fit quadratic curve to these and haven't figured out how to do it correctly  
- The plot that includes all species shows that *once* a given species has ramped up reproduction, the log-log plot of reproductive investment vs final weight is nearly linear across all species; in other words, to first order, all species allocate the same amount to reproduction relative to their weight.
- However, it is important to note that species early in their reproductive trajectory do not follow this pattern
```{r, echo=FALSE,message=FALSE}
plot(ReproInv~FinalWeight, SummaryInd, log="xy",col=col.spp[as.factor(species)],pch=16)
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

plot(ReproInv~FinalWeight, SummaryInd, log="xy", col=col.age[as.factor(age)],pch=16)
legend("bottomright",legend=names(labels.age),col=col.age,pch=16, cex=.8,bty="n")

plot((1+ReproInv_mean)~FinalWeight_mean,SummarySppAge,log="xy", col=col.spp[as.factor(species)],pch=16)
legend("bottomright",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

yvar <- "ReproInv"
xvar <- "FinalWeight"

@@

```{r, echo=FALSE,message=FALSE}
plot(RA~FinalWeight, SummaryInd, log="x", col=col.spp[as.factor(species)],pch=16)
legend("topright",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

plot(RA_mean~FinalWeight_mean,SummarySppAge,log="x", col=col.spp[as.factor(species)],pch=16)
legend("topright",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

yvar <- "RA"
xvar <- "FinalWeight"

@@

These plots show very clearly that for all species RGR suddenly dips when RA becomes non-zero. Plotting against "FinalWeight" or "GrowthInvestment" is much less clear - in other words, an individual's height is the best way to sort them into plants investing more in growth vs reproduction. There are of course some non-reproductive plants with "PropH" above the obvious cut-off, but these are almost certainly "reproductively mature" plants not reproducing in a given year.

>>>>

#RV curves
###Classic RV curves - reproductive investment versus weight
- These are mostly reasonably nice - of course need to fit quadratic curve to these and haven't figured out how to do it correctly  
- The plot that includes all species shows that *once* a given species has ramped up reproduction, the log-log plot of reproductive investment vs final weight is nearly linear across all species; in other words, to first order, all species allocate the same amount to reproduction relative to their weight.
- However, it is important to note that species early in their reproductive trajectory do not follow this pattern
```{r, echo=FALSE,message=FALSE}
plot(ReproInv~total_weight, SummaryInd, log="xy",col=col.spp[as.factor(species)],pch=16)
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

plot(ReproInv~total_weight, SummaryInd, log="xy", col=col.age[as.factor(age)],pch=16)
legend("bottomright",legend=names(labels.age),col=col.age,pch=16, cex=.8,bty="n")

plot((1+ReproInv_mean)~total_weight_mean,SummarySppAge,log="xy", col=col.spp[as.factor(species)],pch=16)
legend("bottomright",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

yvar <- "ReproInv"
xvar <- "total_weight"

@@

```{r, echo=FALSE,message=FALSE}
plot(RA~total_weight, SummaryInd, log="x", col=col.spp[as.factor(species)],pch=16)
legend("topright",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

plot(RA_mean~total_weight_mean,SummarySppAge,log="x", col=col.spp[as.factor(species)],pch=16)
legend("topright",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

yvar <- "RA"
xvar <- "total_weight"

@@

These plots show very clearly that for all species RGR suddenly dips when RA becomes non-zero. Plotting against "total_weight" or "GrowthInvestment" is much less clear - in other words, an individual's height is the best way to sort them into plants investing more in growth vs reproduction. There are of course some non-reproductive plants with "PropH" above the obvious cut-off, but these are almost certainly "reproductively mature" plants not reproducing in a given year.

### 2015-05-20 results_RAschedules_shorter.Rmd
Rename downstream variables

###Plots of size against reproductive investment
* Plots of total reproductive investment against size or height are both highly significant
```{r, echo=FALSE,message=FALSE, fig.width=7, fig.height=7}
mod <- lm(log(FinalWeight)~log(ReproInv), data=subset(SummaryInd,ReproInv>0))
plot(ReproInv~FinalWeight, SummaryInd, log="xy",col=col.spp[as.factor(species)],pch=16)
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")
words.bottom.right.logxy(mod)

@@

```{r, echo=FALSE,message=FALSE, fig.width=7, fig.height=7}
par(mfrow=c(1,1), cex=1, omi=c(.1,.1,.1,.1), mai=c(1.1,1.1,.1,0.2)) 
mod <- lm(log(FinalWeight)~asin(sqrt(RA)), data=subset(SummaryInd,ReproInv>0))
plot(asin(sqrt(RA))~FinalWeight, SummaryInd, log="x",col=col.spp[as.factor(species)],pch=16)
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")
words.top.right.logx(mod)

@@

yvar2 <- "ReproInv"
xvar2 <- "FinalWeight"

@@

* Reproductive investment vs. weight for each species
```{r, echo=FALSE,message=FALSE, fig.width=7, fig.height=7}
yvar <- "ReproInv"
xvar <- "FinalWeight"

>>>>

###Plots of size against reproductive investment
* Plots of total reproductive investment against size or height are both highly significant
```{r, echo=FALSE,message=FALSE, fig.width=7, fig.height=7}
mod <- lm(log(total_weight)~log(ReproInv), data=subset(SummaryInd,ReproInv>0))
plot(ReproInv~total_weight, SummaryInd, log="xy",col=col.spp[as.factor(species)],pch=16)
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")
words.bottom.right.logxy(mod)

@@

```{r, echo=FALSE,message=FALSE, fig.width=7, fig.height=7}
par(mfrow=c(1,1), cex=1, omi=c(.1,.1,.1,.1), mai=c(1.1,1.1,.1,0.2)) 
mod <- lm(log(total_weight)~asin(sqrt(RA)), data=subset(SummaryInd,ReproInv>0))
plot(asin(sqrt(RA))~total_weight, SummaryInd, log="x",col=col.spp[as.factor(species)],pch=16)
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")
words.top.right.logx(mod)

@@

yvar2 <- "ReproInv"
xvar2 <- "total_weight"

@@

* Reproductive investment vs. weight for each species
```{r, echo=FALSE,message=FALSE, fig.width=7, fig.height=7}
yvar <- "ReproInv"
xvar <- "total_weight"

### 2015-05-20 results_accessory_shorter.Rmd
Rename downstream variables

```{r, echo=FALSE,message=FALSE,,fig.width=14, fig.height=14}
par(mfrow=c(2,2), cex=1, omi=c(2,0.3,.5,0.3), mai=c(.5,1,.5,0.3))
mod <- lm (log(FinalWeight)~log(ReproInv), data=subset(SummaryInd,ReproInv>0))
plot(ReproInv~FinalWeight,data=subset(SummaryInd,ReproInv>0), log="xy",xlab="", col=col.spp[as.factor(species)], pch=16, ylab="Total reproductive investment (mg)",xlim=c(100,2000000))
words.bottom.right.logxy(mod)

mod <- lm (log(FinalWeight)~log(fruit_weight), data=subset(SummaryInd,fruit_weight>0))
plot(fruit_weight~FinalWeight,data=subset(SummaryInd,fruit_weight>0), log="xy",col=col.spp[as.factor(species)], pch=16, ylab="Fruit weight (mg)",xlim=c(100,2000000),xlab="")
words.bottom.right.logxy(mod)

mod <- lm (log(FinalWeight)~log(Prop), data=subset(SummaryInd,Prop>0))
plot(Prop~FinalWeight,data=subset(SummaryInd,Prop>0), log="xy",col=col.spp[as.factor(species)], pch=16, xlab="", ylab="Propagule investment (mg)", xlim=c(100,2000000))
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")
words.bottom.right.logxy(mod)

mod <- lm (log(FinalWeight)~log(PerPropagule), data=subset(SummaryInd,PerPropagule>0))
plot(PerPropagule~FinalWeight,data=subset(SummaryInd,PerPropagule>0), log="xy",col=col.spp[as.factor(species)], pch=16, ylab="Proportion reproductive investment to propagules",xlim=c(100,2000000),xlab="Total size (mg)")
words.bottom.right.logxy(mod)
mtext("Total size (mg)", 1, outer=TRUE)
```

@@

mod <- lm(log(FinalWeight_mean)~asin(sqrt(PerAcc_mean)),SummarySpp)
plot(asin(sqrt(PerAcc_mean))~FinalWeight_mean,SummarySpp, log="x", pch=16,col=col.spp[as.factor(species)],ylab="",xlab="Total weight (mg)",cex=1.3)
words.bottom.right.logx(mod)

>>>>

```{r, echo=FALSE,message=FALSE,,fig.width=14, fig.height=14}
par(mfrow=c(2,2), cex=1, omi=c(2,0.3,.5,0.3), mai=c(.5,1,.5,0.3))
mod <- lm (log(total_weight)~log(ReproInv), data=subset(SummaryInd,ReproInv>0))
plot(ReproInv~total_weight,data=subset(SummaryInd,ReproInv>0), log="xy",xlab="", col=col.spp[as.factor(species)], pch=16, ylab="Total reproductive investment (mg)",xlim=c(100,2000000))
words.bottom.right.logxy(mod)

mod <- lm (log(total_weight)~log(fruit_weight), data=subset(SummaryInd,fruit_weight>0))
plot(fruit_weight~total_weight,data=subset(SummaryInd,fruit_weight>0), log="xy",col=col.spp[as.factor(species)], pch=16, ylab="Fruit weight (mg)",xlim=c(100,2000000),xlab="")
words.bottom.right.logxy(mod)

mod <- lm (log(total_weight)~log(Prop), data=subset(SummaryInd,Prop>0))
plot(Prop~total_weight,data=subset(SummaryInd,Prop>0), log="xy",col=col.spp[as.factor(species)], pch=16, xlab="", ylab="Propagule investment (mg)", xlim=c(100,2000000))
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")
words.bottom.right.logxy(mod)

mod <- lm (log(total_weight)~log(PerPropagule), data=subset(SummaryInd,PerPropagule>0))
plot(PerPropagule~total_weight,data=subset(SummaryInd,PerPropagule>0), log="xy",col=col.spp[as.factor(species)], pch=16, ylab="Proportion reproductive investment to propagules",xlim=c(100,2000000),xlab="Total size (mg)")
words.bottom.right.logxy(mod)
mtext("Total size (mg)", 1, outer=TRUE)
```

@@

mod <- lm(log(total_weight_mean)~asin(sqrt(PerAcc_mean)),SummarySpp)
plot(asin(sqrt(PerAcc_mean))~total_weight_mean,SummarySpp, log="x", pch=16,col=col.spp[as.factor(species)],ylab="",xlab="Total weight (mg)",cex=1.3)
words.bottom.right.logx(mod)

### 2015-05-20 results_accessory_tissues.Rmd
Rename downstream variables

##Do individual species decrease accessory costs as they increase in size, age or RA?
###Change in accesory costs With size:  
- PerAcc is arcsine transformed, size is log transformed  
- COER, HATE, and PILI show a significant decrease in accessory costs with increasing size (p<0.05)  
- GRSP shows a marginally significant decrease in accessory costs with increasing size (0.05<p<0.10)  
- PEPU shows a marginally significant INCREASE in accessory costs with increasing size (0.05<p<0.10)  
```{r, echo=FALSE, message=FALSE}
yvar <- "PerAcc"
xvar <- "FinalWeight"

@@

##Next considering if the amount of energy spent on aborted buds shifts with age, size, RA  
###Change in aborted bud investment with size:  
- PerPrePol_A is arcsine transformed, size is log transformed  
- GRSP, HATE, HEPU show a significant decrease in aborted buds with increasing size (p<0.05)  
- PUTU shows a marginally significant decrease in accessory costs with increasing RA (p ~ 0.10)  
```{r, echo=FALSE, message=FALSE}
yvar <- "PerPrePol_A"
xvar <- "FinalWeight"

@@

```{r, echo=FALSE, message=FALSE}
yvar <- "PerPrePol_All"
xvar <- "FinalWeight"

>>>>

##Do individual species decrease accessory costs as they increase in size, age or RA?
###Change in accesory costs With size:  
- PerAcc is arcsine transformed, size is log transformed  
- COER, HATE, and PILI show a significant decrease in accessory costs with increasing size (p<0.05)  
- GRSP shows a marginally significant decrease in accessory costs with increasing size (0.05<p<0.10)  
- PEPU shows a marginally significant INCREASE in accessory costs with increasing size (0.05<p<0.10)  
```{r, echo=FALSE, message=FALSE}
yvar <- "PerAcc"
xvar <- "total_weight"

@@

##Next considering if the amount of energy spent on aborted buds shifts with age, size, RA  
###Change in aborted bud investment with size:  
- PerPrePol_A is arcsine transformed, size is log transformed  
- GRSP, HATE, HEPU show a significant decrease in aborted buds with increasing size (p<0.05)  
- PUTU shows a marginally significant decrease in accessory costs with increasing RA (p ~ 0.10)  
```{r, echo=FALSE, message=FALSE}
yvar <- "PerPrePol_A"
xvar <- "total_weight"

@@

```{r, echo=FALSE, message=FALSE}
yvar <- "PerPrePol_All"
xvar <- "total_weight"
